---
title: "KOGO_LCB_pipeline"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

## **Load package and dataset**
SingleCellExperiment object which quality control (QC) was done should be loaded.
```{r}
library(scran)
library(scater)
```

```{r, echo=F}
setwd('D:/KOGO/dataset/KOGO2022/')

sce <- readRDS('./sce.rds')
```

```{r}
sce
```
## **Normalization**
To remove cell-specific biases, cells are clustered using **quickCluster()** and cell-specific size factors are calculated using **computeSumFactors()** of scran R package. Raw counts of each cell are divided by cell-specific size factor and log2-transformed with a pseudocount of 1.
```{r}
clusters <- quickCluster(sce)
sce <- computeSumFactors(sce, clusters = clusters)
sce.norm <- logNormCounts(sce,  pseudo_count = 1)
```

## **Feature selection(calculate highly variable genes)**
To find genes contain useful information about the biology of the data, highly variable genes (HVGs) are defined by selecting the most variable genes based on their expression across cells. Genes with < 0.05 of false discovery rate (FDR) are identified as HVGs.
```{r}
gene_variance <- modelGeneVar(sce.norm)
```


```{r, echo=FALSE}
plot(gene_variance$mean, gene_variance$total,
     xlab = "mean_log-expression",
     ylab = "variance")
curve(metadata(gene_variance)$trend(x),
      col = "blue",
      add = T)
```

```{r}
hvg.norm <- getTopHVGs(gene_variance, fdr.threshold = 0.05)
length(hvg.norm) 
```


## **Create seurat object**
For downstream analysis, we create Seurat object containing raw and normalized gene-by-cell count matrices. Column data (cell information) is preserved. Normalized data is scaled and principal components (PCs) are calculated by a gene-by-cell matrix with HVGs.
```{r}
library(Seurat)

seurat <- as.Seurat(sce.norm,
                    counts = "counts",
                    data = "logcounts")
VariableFeatures(seurat) <- hvg.norm
```

## **Dimensionality reduction**
Normalized data is scaled and principal components (PCs) are calculated by a gene-by-cell matrix with HVGs.
```{r}
all.genes <- rownames(seurat)
seurat <- ScaleData(seurat, features = all.genes)

seurat <- RunPCA(seurat, features = VariableFeatures(seurat))
```

```{r}
ElbowPlot(seurat, ndims = 50)
```

We set to 15 PCs for clustering and visualization. After clustering and visualization, cells are plotted in the two-dimensional TSNE or UMAP plot and cell information can be also shown.
```{r}
PCs <- 15

seurat <- FindNeighbors(seurat,
                        dims = 1:PCs)
seurat <- FindClusters(seurat,
                       resolution = 0.2)
seurat <- RunUMAP(seurat, dims = 1:PCs)
seurat <- RunTSNE(seurat, dims = 1:PCs)

```

```{r, echo=FALSE}
DimPlot(seurat, label=TRUE)
```

## **Canonical cell type marker gene expression**
Based on 'seurat_clusters' after clustering, cells are grouped according to their cell types as annotated based on known cell lineage-specific marker genes
```{r}
marker.genes<- list(T.cell = c('CD3D','CD3E','TRAC'),
                    Monocyte = c('CD14','S100A8','FCGR3A'),
                    NK.cell = c('NCAM1','NKG7','TRDC'),
                    B.cell = c('CD79A','CD83','MS4A1'),
                    Classical.Dendritic = c('CLEC9A','FCER1A','CLEC10A'),
                    Plasmacytoid.Dendritic = c('CLEC4C','IL3RA','LILRA4'),
                    Plasma.cell = c('JCHAIN','TNFRSF17','SDC1'))
```

### T.cell
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$T.cell, order = T, ncol = 3)
```

### Monocyte
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$Monocyte, order = T, ncol = 3)
```

### NK.cell
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$NK.cell, order = T, ncol = 3)
```

### B.cell
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$B.cell, order = T, ncol = 3)
```

### Classical.Dendritic
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$Classical.Dendritic, order = T, ncol = 3)
```

### Plasmacytoid.Dendritic
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$Plasmacytoid.Dendritic, order = T, ncol = 3)
```

### Plasma.cell
```{r, fig.height = 4, fig.width = 12}
FeaturePlot(seurat, features = marker.genes$Plasma.cell, order = T, ncol = 3)
```

## Marker gene expression on Heatmap

```{r}
library(pheatmap)
library(RColorBrewer)

avgExprs <- AverageExpression(seurat,
                              features = unlist(marker.genes),
                              assays = "originalexp", slot = "data")

scaledExprs <- t(scale(t(avgExprs$originalexp)))
scaledExprs[scaledExprs > -min(scaledExprs)] <- -min(scaledExprs)

palette_length = 100
my_color = my_color <- colorRampPalette(rev(brewer.pal(11, "RdBu")))(palette_length)

my_breaks <- c(seq(min(scaledExprs), 0,
                   length.out=ceiling(palette_length/2) + 1),
               seq(max(scaledExprs)/palette_length,
                   max(scaledExprs),
                   length.out=floor(palette_length/2)))

pheatmap(scaledExprs,
         cluster_cols = T, cluster_rows = F, clustering_method = "ward.D2",
         treeheight_col = 0,
         breaks = my_breaks, color=my_color,
         labels_row = as.expression(lapply(rownames(scaledExprs), function(a) bquote(italic(.(a))))),
         angle_col = 315)

```

## **Cell type annotation**

```{r}
seurat$celltype = as.character(seurat$seurat_clusters)

seurat$celltype[seurat$celltype %in% c('1')] <- 'NK.cell'
seurat$celltype[seurat$celltype %in% c('0','4','6')] <- 'T.cell'
seurat$celltype[seurat$celltype %in% c('3')] <- 'B.cell'
seurat$celltype[seurat$celltype %in% c('8')] <- 'Plasma.cell'
seurat$celltype[seurat$celltype %in% c('2','5')] <- 'Monocyte'
seurat$celltype[seurat$celltype %in% c('7')] <- 'Classical.Dendritic'
seurat$celltype[seurat$celltype %in% c('9')] <- 'Plasmacytoid.Dendritic'

```

```{r, echo=FALSE}
DimPlot(seurat, group.by = 'celltype', label=TRUE)
```

## **Reference**

Pekayvaz, K., Leunig, A., Kaiser, R. et al. Protective immune trajectories in early viral containment of non-pneumonic SARS-CoV-2 infection. Nat Commun 13, 1018 (2022).

McCarthy, D. J., Campbell, K. R., Lun, A. T. & Wills, Q. F. Scater: pre-processing, quality control, normalization and visualization of single-cell RNA-seq data in R. Bioinformatics 33, 1179–1186 (2017)

Lun, A. T., McCarthy, D. J. & Marioni, J. C. A step-by-step workflow for low-level analysis of single-cell RNA-seq data with Bioconductor. F1000Res 5, 2122 (2016).

Yuhan Hao, Stephanie Hao, Erica Andersen-Nissen, William M. Mauck, Shiwei Zheng, Andrew Butler, Maddie J. Lee, Aaron J. Wilk, Charlotte Darby, Michael Zager, Paul Hoffman, Marlon Stoeckius, Efthymia Papalexi, Eleni P. Mimitou, Jaison Jain, Avi Srivastava, Tim Stuart, Lamar M. Fleming, Bertrand Yeung, Angela J. Rogers, Juliana M. McElrath, Catherine A. Blish, Raphael Gottardo, Peter Smibert, Rahul Satija,
Integrated analysis of multimodal single-cell data, Cell,184(13):3573-3587(2021).